\newpage
\addcontentsline{toc}{chapter}{Acknowledgements}
\section*{Acknowledgements}
This thesis concludes my master's degree in Environment and Natural Resources at the Norwegian University of Life Sciences (NMBU). The work was conducted in collaboration with the Norwegian Geotechnical Institute (NGI) as part of the Valorization of Organic Waste (VOW) project funded by The Research Council of Norway (NFR 299070) with contributions from the SLUDGEFFECT project (NFR 302371). 

I ow great thanks to all the people who have supported and assisted me, making the completion of this work possible. I would like to express my deepest gratitude to all the amazing employees at NGI who have supported let me conduct all of my tests in their environmental chemistry laboratory.
my supervisors, Erlend Sørmo and Gerard Cornelissen for being constant sources of knowledge and support. 
  

Thank you to Lindum AS and all the workers there  who have included me in 
Gabriela Castro Varela for sample processing at NTNU with SPE and LC-MS/MS

Michel Hubert (PhD candidate at NGI), I thank you for all your great ideas and discussion about common issues and ideas when writing our theses. You have brought much joy to long days of writing. 

Hans Peter Arp for help with data interpretation and presentation in the later stages of the work. 

At last, I would like to thank my family and class mates for their support. My diet would not been as healthy without your support with cooking during the final stages of writing this thesis. 

I would like to express my deepest gratitude to Aurora Hofman and Raol Wolf for helping me with coding and plotting in R. You have saved my days from endless frustration. My thanks also go to Egil, Marit and Christian from NGI who introduced me to \LaTeX and helped me with the formatting of this thesis. A great thanks to John for proofreading my thesis. 
